---
title: "GIS Tutorial HW"
author: "Jordan Daley"
date: "2023-04-11"
output: html_document
---
```{r}
library("tmaptools")
library("sf")
library("leaflet")
```

```{r}
install.packages("scales")
library("scales")
```

```{r}
library(tidyverse)
library(tmap)
library(tmaptools)
library(leaflet)
library(sf)
library(leaflet.extras)
library(dplyr)
library(rio)
library(sp)
```

##  Step 1: Election result data
```{r}
nhdatafile <- "data/NHD2016.xlsx"
```

## # Step 2: Add columns for percents and margins
```{r}
nhdata$SandersMarginVotes <- nhdata$Sanders - nhdata$Clinton
nhdata$SandersPct <- (nhdata$Sanders - nhdata$Clinton) / (nhdata$Sanders + nhdata$Clinton) # Will use formatting later to multiply by a hundred 
nhdata$ClintonPct <- (nhdata$Clinton - nhdata$Sanders) / (nhdata$Sanders + nhdata$Clinton)
nhdata$SandersMarginPctgPoints <- nhdata$SandersPct - nhdata$ClintonPct
```

```{r}
usshapefile <- "data/cb_2014_us_county_5m/cb_2014_us_county_5m.shp"
usgeo <- read_shape(file=usshapefile, as.sf = TRUE)
nhgeo <- usgeo[usgeo@data$STATEFP=="33",]
nhgeo <- filter(usgeo, STATEFP=="33")
```

## Step 4: Merge spatial and results data
```{r}
str(nhgeo$NAME)
# Factor w/ 1921 levels "Abbeville","Acadia",..: 1470 684 416 1653 138 282 1131 1657 334 791

str(nhdata$County)
# chr [1:11] "Belknap" "Carroll" "Cheshire" "Coos" "Grafton"

nhgeo$NAME <- as.character(nhgeo$NAME)

nhgeo <- nhgeo[order(nhgeo$NAME),]
nhdata <- nhdata[order(nhdata$County),]

identical(nhgeo$NAME,nhdata$County )
[1] TRUE

nhmap <- append_data(nhgeo, nhdata, key.shp = "NAME", key.data="County")

str(nhmap)
```

## Create a static map
```{r}
qtm(nhmap, "SandersMarginVotes")

qtm(nhmap, "SandersMarginPctgPoints")

tm_shape(nhmap) +
tm_fill("SandersMarginVotes", title="Sanders Margin, Total Votes", palette = "PRGn") +
tm_borders(alpha=.5) +
tm_text("NAME", size=0.8)
tm_style_classic()


```

## Create palette and pop-ups for interactive 
```{r}
mypalette <- colorFunction(palette = "colors I want", domain = mydataframe$dataColumnToMap)

clintonPalette <- colorNumeric(palette = "Blues", domain=nhmap$ClintonPct)

nhpopup <- paste0("County: ", nhmap$County,
"Sanders ", percent(nhmap$SandersPct), " - Clinton ", percent(nhmap$ClintonPct))

nhmap <- rename(nhmap, County = NAME)
```

## Step 7: Generate an interactive map
```{r}
leaflet(nhmap) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(stroke=FALSE, 
              smoothFactor = 0.2,
              fillOpacity = .8, 
              popup=nhpopup,
              color= ~clintonPalette(nhmap$ClintonPct)

  nhmap_projected <- sf::st_transform(nhmap, "+proj=longlat +datum=WGS84")
```

## Step 8: Add palettes for multi-layer map
```{r}
minpct <- min(c(scmap$Donald.J.TrumpPct, scmap$Marco.RubioPct , scmap$Ted.CruzPct))
maxpct <- max(c(scmap$Donald.J.TrumpPct, scmap$Marco.RubioPct , scmap$Ted.CruzPct))

trumpPalette <- colorNumeric(palette = "Purples", domain=c(minpct, maxpct))
rubioPalette <- colorNumeric(palette = "Reds", domain = c(minpct, maxpct))
cruzPalette <- colorNumeric(palette = "Oranges", domain = c(minpct, maxpct))
```

```{r}
winnerPalette <- colorFactor(palette=c("#984ea3", "#e41a1c"), domain = scmap$winner)
edPalette <- colorNumeric(palette = "Blues", domain=scmap$PctCollegeDegree)
```

```{r}
scpopup <- paste0("County: ", scmap@data$County,
"Winner: ", scmap@data$winner,
"Trump: ", percent(scmap$Donald.J.TrumpPct),
"Rubio: ", percent(scmap$Marco.RubioPct),
"Cruz: ", percent(scmap$Ted.CruzPct),
"Pct w college ed: ", scmap$PctCollegeDegree, "% vs state-wide avg of 25%")
```

```{r}
leaflet(scmap) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(stroke=TRUE,
              weight=1,
              smoothFactor = 0.2,
              fillOpacity = .75,
              popup=scpopup, 
              color= ~winnerPalette(scmap$winner),
              group="Winners"
 ) %>%
    addLegend(position="bottomleft", colors=c("#984ea3", "#e41a1c"), labels=c("Trump", "Rubio"))
```

## Step 9: 
```{r}
scGOPmap <- leaflet(scmap) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(stroke=TRUE,
              weight=1,
              smoothFactor = 0.2,
              fillOpacity = .75,
              popup=scpopup, 
              color= ~winnerPalette(scmap$winner),
              group="Winners"
  ) %>% 
    addLegend(position="bottomleft", colors=c("#984ea3", "#e41a1c"), labels=c("Trump", "Rubio"))  %>%

  addPolygons(stroke=TRUE,
     weight=1,
     smoothFactor = 0.2, 
     fillOpacity = .75, 
     popup=scpopup, 
     color= ~trumpPalette(scmap$Donald.J.TrumpPct),
     group="Trump"
    ) %>%

  addPolygons(stroke=TRUE,
              weight=1,
              smoothFactor = 0.2, 
              fillOpacity = .75, 
              popup=scpopup, 
              color= ~rubioPalette(scmap$Marco.RubioPct),
              group="Rubio"
  ) %>%

  addPolygons(stroke=TRUE,
              weight=1,
              smoothFactor = 0.2, 
              fillOpacity = .75, 
              popup=scpopup, 
              color= ~cruzPalette(scmap$Ted.CruzPct),
              group="Cruz"
  ) %>%

  addPolygons(stroke=TRUE,
              weight=1,
              smoothFactor = 0.2, 
              fillOpacity = .75, 
              popup=scpopup, 
              color= ~edPalette(scmap$PctCollegeDegree),
              group="College degs"
  ) %>%

  addLayersControl(
      baseGroups=c("Winners", "Trump", "Rubio", "Cruz", "College degs"),
      position = "bottomleft",
      options = layersControlOptions(collapsed = FALSE)
      ) 
```

## Step 10
```{r}
# install.packages("htmlwidgets")
save(widget=scGOPmap2, file="scGOPprimary_withdependencies.html", selfcontained=FALSE, libdir = "js")
```


```{r}
# install.packages("htmlwidgets")
save(widget=scGOPmap2, file="scGOPprimary_withdependencies.html", selfcontained=FALSE, libdir = "js")